Report

### 2022-07-24 Shiny/Test/Dashboard downloads test.Rmd
Experiments with generating a downloadable report.

@@

- **Annual result**: Based on the entry *Årsresultat* in the Proff.no data.
- **Balance**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
- **Revenues**: Based on the category *Revenue* in the Excel sheet.
- **Expenses**: Based on the category *Expenses* in the Excel sheet.
- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
  - A common measure of transparency, but should normally be evaluated in context.

```{r value_boxes, include=FALSE}
# Shiny actions
observe({ # wait for button click

@@

  ## Public/private proportion--------------------------------------------------
  # Creating public/private proportion
  KPI_proportion <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        transmute(
          proportion = sum(revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) / sum(revenue_public_grants, revenue_private_contributions, revenue_membership_fees, na.rm = TRUE) * 100
        ) |>
        pull() |>
        first()
    }
  })

@@

  ## Annual result--------------------------------------------------------------
  arsresultat_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        ggplot(aes(year, arsresultat, fill = annual_profit)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$annual_profit) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Annual result (in thousands)"
        )
    }
  })

@@

  ## Balance--------------------------------------------------------------------
  balance_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        ggplot(aes(year, sum_egenkapital_og_gjeld, fill = balance_positive)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$balance_positive) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Balance (in thousands)",
        )
    }
  })

>>>>

If desired, a copy of the analysis in html format may then be downloaded.

```{r sidebar_button}
# download
uiOutput("downloadUI")
```

@@

- **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
- **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
- **Revenues**: Based on the category *Revenue* in the Excel sheet.
- **Expenses**: Based on the category *Expenses* in the Excel sheet.
- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
  - A common measure of transparency, but should normally be evaluated in context.

```{r value_boxes, include=FALSE}
# Shiny actions

  # Report Output ==============================================================

  output$downloadUI <- renderUI( {
  downloadButton("report", "Generate report")
})

  output$report <- downloadHandler(
    filename = "report.html",
    content = function(file) {
      tempReport <- file.path(tempdir(), "report.Rmd")
      file.copy("report.Rmd", tempReport, overwrite = TRUE)

      params <- list(htmlfile = input$html$datapath,
                     excelfile = input$excel$datapath,
                     start = input$start,
                     end = input$end)

      rmarkdown::render(tempReport, output_file = file,
                        params = params,
                        envir = new.env(parent = globalenv()))
    }
  )

observe({ # wait for button click

@@

  ## Public/private proportion--------------------------------------------------
  # Creating public/private proportion
  KPI_proportion <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        transmute(
          proportion =
            sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) /
              sum(revenue_public_grants, revenue_private_contributions,
                revenue_membership_fees,
                na.rm = TRUE
              ) * 100
        ) |>
        pull() |>
        first()
    }
  })

@@

  ## Annual result--------------------------------------------------------------
  arsresultat_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        # renaming for plotly hover
        rename(
          Year = year,
          `Net income` = arsresultat,
          `Net positive` = annual_profit
        ) |> 
        ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$annual_profit) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Net income (in thousands)"
        )
    }
  })

@@

  ## Balance--------------------------------------------------------------------
  balance_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        # rename for plotly hover
        rename(
          Year = year,
          `Total equity/liabilities` = sum_egenkapital_og_gjeld,
          `Positive equity` = balance_positive
        ) |> 
        ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$balance_positive) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Total equity/liabilities (in thousands)",
        )
    }
  })

### 2022-07-24 Shiny/Test/report.Rmd
Experiments with generating a downloadable report.

---
title: "NGO analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
runtime: shiny
---

```{r setup, include=FALSE}
# Loading packages using pacman to ensure they're all installed
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
# Using the fontawesome package for more up-to-date (and currently supported)
# Fontawesome icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

Sidebar {.sidebar}
=======================================================================

### File upload

Please upload an html copy of annual accounts from [Proff.no](https://www.proff.no) and enter the start and end years from that table.

Then upload an Excel file in the set format.

Click 'Analyse' to analyse the data.

```{r sidebar}
# HTML
fileInput("html", label = "Upload html file", accept = ".html")

# Year inputs
numericInput("start", label = "Start", value = 2002)
numericInput("end", label = "End", value = 2021)

# Excel
fileInput("excel", "Upload Excel file", accept = ".xlsx")

# Button
actionButton("click", "Analyse")
```

If desired, a copy of the analysis in html format may then be downloaded.

```{r sidebar_button}
# download
uiOutput("downloadUI")
```

Dashboard
=======================================================================

Row 
-----------------------------------------------------------------------

### KPI Fundraising
```{r}
valueBoxOutput("KPI_box_fundraising")
```

### KPI OPX
```{r}
valueBoxOutput("KPI_box_OPX")
```

### KPI Private growth
```{r}
valueBoxOutput("KPI_box_private")
```

### KPI Public growth
```{r}
valueBoxOutput("KPI_box_public")
```

### KPI Private/public proportion
```{r}
valueBoxOutput("KPI_box_proportion")
```

Row
-----------------------------------------------------------------------

### Annual result
```{r}
plotlyOutput("arsresultat_plotly")
```

### Balance
```{r}
plotlyOutput("balance_plotly")
```

Row 
-----------------------------------------------------------------------

### Revenues
```{r}
plotlyOutput("revenue_plotly")
```

### Expenses
```{r}
plotlyOutput("expenses_plotly")
```

Row 
-----------------------------------------------------------------------

### Transparency: CEO compensation
```{r}
plotlyOutput("CEO_plotly")
```

Details
=======================================================================

Row 
-----------------------------------------------------------------------

### Key Performance Indicators

Row 
-----------------------------------------------------------------------

### Plots

The dashboard includes five plots, three based on data from [Proff.no](https://www.proff.no) and two based on data from the Excel sheet.

- **Net income/loss**: Based on the entry *Årsresultat* in the Proff.no data.
- **Total equity/liabilities**: Based on the entry *Sum egenkapital og gjeld* in the Proff.no data.
- **Revenues**: Based on the category *Revenue* in the Excel sheet.
- **Expenses**: Based on the category *Expenses* in the Excel sheet.
- **CEO compensation**: Based on the category *Lederlønn* in the Proff.no data. Not in thousands. Includes the entries *Lønn* and *Leder annen godtgjørelse*.
  - A common measure of transparency, but should normally be evaluated in context.

```{r value_boxes, include=FALSE}
# Shiny actions

  # Report Output ==============================================================

  output$downloadUI <- renderUI( {
  downloadButton("report", "Generate report")
})

  output$report <- downloadHandler(
    filename = "report.html",
    content = function(file) {
      tempReport <- file.path(tempdir(), "report.Rmd")
      file.copy("report.Rmd", tempReport, overwrite = TRUE)

      params <- list(htmlfile = input$html$datapath,
                     excelfile = input$excel$datapath,
                     start = input$start,
                     end = input$end)

      rmarkdown::render(tempReport, output_file = file,
                        params = params,
                        envir = new.env(parent = globalenv()))
    }
  )

observe({ # wait for button click

  # Data frames ================================================================

  # creating data frame from html file
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))
    widen_excel(input$excel$datapath)
  })

  # Value Boxes ================================================================

  ## Fundraising ROI------------------------------------------------------------
  # Creating Fundraising ROI value
  KPI_fundraising_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # most recent year
        filter(year == max(year)) |>
        transmute(
          # Calculating the value as a percentage; removing all other
          KPI_fundraising_num =
            (sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) - expenses_fundraising) /
              expenses_fundraising * 100
        ) |>
        # just want the single number
        pull() |>
        first()
    }
  })

  # Creating value box for fundraising ROI
  # Unsure about colour levels
  output$KPI_box_fundraising <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_fundraising_num()),
      caption = "Fundraising ROI",
      # Colour depends on value
      color = if (is.na(KPI_fundraising_num())) {
        "primary"
      } else if (KPI_fundraising_num() < 100) {
        "danger"
      } else if (KPI_fundraising_num() < 200) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-donate"
    )
  })

  ## OPX KPI--------------------------------------------------------------------
  # Creating OPX KPI
  KPI_OPX_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # most recent year
        filter(year == max(year)) |>
        # Calculating the value as a %
        transmute(
          KPI_OPX =
            sum(revenue_private_contributions, revenue_investments, assets_cash_in_bank_and_cash_equivalents, na.rm = TRUE) /
              sum(liability_short_term_grants_payable, liability_revocable_endowments, liability_long_term_grants_payable, na.rm = TRUE) * 100
        ) |>
        # Just the single number
        pull() |>
        first()
    }
  })

  # Creating value box for OPX KPI
  output$KPI_box_OPX <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_OPX_num()),
      caption = "Liquid assets / obligations",
      color = if (is.na(KPI_OPX_num())) {
        "primary"
      } else if (KPI_OPX_num() < 100) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-seedling"
    )
  })

  ## Private donations growth---------------------------------------------------
  # Creating private donations growth
  KPI_private_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # Calculating the two most recent years as a %
        transmute(
          growth =
            (revenue_private_contributions[1] -
              revenue_private_contributions[2]) /
              revenue_private_contributions[2] * 100
        ) |>
        # Pulling the first instance of the number
        pull() |>
        first()
    }
  })

  # Creating value box for private donations growth
  output$KPI_box_private <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_private_num()),
      caption = "Private donation growth",
      color = if (is.na(KPI_private_num())) {
        "primary"
      } else if (KPI_private_num() < 0) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-hands-helping"
    )
  })

  ## Public grants growth-------------------------------------------------------
  # Creating public grants growth
  KPI_public_num <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        # Calculating the two most recent years as a %
        transmute(
          growth =
            (revenue_public_grants[1] - revenue_public_grants[2]) /
              revenue_public_grants[2] * 100
        ) |>
        # Pulling the first instance of the number
        pull() |>
        first()
    }
  })

  # Creating value box for public grants growth
  output$KPI_box_public <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_public_num()),
      caption = "Public grants growth",
      color = if (is.na(KPI_public_num())) {
        "primary"
      } else if (KPI_public_num() < 0) {
        "warning"
      } else {
        "success"
      },
      icon = "fas fa-hand-holding-usd"
    )
  })

  ## Public/private proportion--------------------------------------------------
  # Creating public/private proportion
  KPI_proportion <- reactive({
    # testing for data quality, returning NA if bad
    if (datatest() == TRUE) {
      NA
    } else {
      exceldata() |>
        transmute(
          proportion =
            sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) /
              sum(revenue_public_grants, revenue_private_contributions,
                revenue_membership_fees,
                na.rm = TRUE
              ) * 100
        ) |>
        pull() |>
        first()
    }
  })

  # Creating value box for private/public proportion
  output$KPI_box_proportion <- renderValueBox({
    valueBox(
      value = KPI_to_string(KPI_proportion()),
      caption = "Private donations",
      # Adding in the same check as elsewhere - colours match if all are NA
      color = if (is.na(KPI_proportion())) {
        "primary"
      } else {
        "info"
      },
      icon = "fas fa-percentage"
    )
  })

  ## Excel data test------------------------------------------------------------
  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
}) |> bindEvent(input$click)
```

```{r plots, include=FALSE}
# It's necessary to split the observers into separate parts, with all necessary
# code duplicated.

observe({

  # Data frames ================================================================

  # creating data frame from html file
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))
    widen_excel(input$excel$datapath)
  })

  # Plots ======================================================================

  ## Annual result--------------------------------------------------------------
  arsresultat_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$arsresultat)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        # renaming for plotly hover
        rename(
          Year = year,
          `Net income` = arsresultat,
          `Net positive` = annual_profit
        ) |> 
        ggplot(aes(Year, `Net income`, fill = `Net positive`)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$annual_profit) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Net income (in thousands)"
        )
    }
  })

  # Creating plotly output
  output$arsresultat_plotly <- renderPlotly(
    ggplotly(arsresultat_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  ## Balance--------------------------------------------------------------------
  balance_plot <- reactive({
    # testing for the presence of data
    if (sum(!is.na(htmldata()$sum_egenkapital_og_gjeld)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      htmldata() |>
        # rename for plotly hover
        rename(
          Year = year,
          `Total equity/liabilities` = sum_egenkapital_og_gjeld,
          `Positive equity` = balance_positive
        ) |> 
        ggplot(aes(Year, `Total equity/liabilities`, fill = `Positive equity`)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # using a function to ensure uniform colours irrespective of data
        scale_fill_green(htmldata()$balance_positive) +
        theme(legend.position = "none") +
        labs(
          x = "", y = "", title = "Total equity/liabilities (in thousands)",
        )
    }
  })

  # Creating plotly output
  output$balance_plotly <- renderPlotly(
    ggplotly(balance_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  ## Expenses-------------------------------------------------------------------
  expenses_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          Other = expenses_other,
          Administrative = expenses_administrative,
          Fundraising = expenses_fundraising,
          `Program services` = expenses_program_services
        ) |>
        # Pivoting to use fill
        pivot_longer(c(Other:`Program services`), names_to = "Expenses") |>
        # Ordering for the plot
        mutate(
          Expenses = ordered(Expenses,
            levels =
              c("Program services", "Other", "Fundraising", "Administrative")
          )
        ) |>
        ggplot(aes(Year, value, fill = Expenses)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        scale_fill_met_d("Hokusai2") +
        labs(x = "", y = "", title = "Expenses (in thousands)")
    }
  })

  # Creating plotly output
  output$expenses_plotly <- renderPlotly(
    ggplotly(expenses_plot())
  )

  ## Revenue--------------------------------------------------------------------
  revenue_plot <- reactive({
    # testing for data quality
    if (datatest() == TRUE) {
      # printing 'bad data' plot if data are bad
      bad_data()
    } else {
      exceldata() |>
        # Renaming for the purposes of the plot + dropping other variables
        transmute(
          Year = as.factor(year),
          `Private contributions` = revenue_private_contributions,
          `Membership fees` = revenue_membership_fees,
          `Public grants` = revenue_public_grants,
          Investments = revenue_investments,
          `Operational activities` = revenue_operational_activities,
          `Commercial activities` = revenue_commercial_activities,
          Other = revenue_other
        ) |>
        # Pivoting to use fill
        pivot_longer(
          c(`Private contributions`:Other),
          names_to = "Revenue source"
        ) |>
        # Ordering for the plot
        mutate(
          `Revenue source` =
            ordered(`Revenue source`,
              levels =
                c(
                  "Other", "Commercial activities",
                  "Operational activities", "Investments",
                  "Membership fees", "Private contributions",
                  "Public grants"
                )
            )
        ) |>
        ggplot(aes(Year, value, fill = `Revenue source`)) +
        geom_col() +
        scale_y_continuous(
          # Excel data are assumed to be in NOK; if this changes, code based on
          # the currency_code() function may be adapted
          labels = label_dollar(prefix = "NOK "),
          # Basing lower limit on the data (not suppressing erroneous negative
          # numbers), and the upper limit on the maximum value in revenues and
          # expenses, to match the two plots
          limits = c(NA, maximum())
        ) +
        # Specifying the number of values, as the base palette only includes 6
        scale_fill_manual(values = met.brewer("Hokusai2", 7)) +
        labs(x = "", y = "", title = "Revenue (in thousands)")
    }
  })

  # Creating plotly output
  output$revenue_plotly <- renderPlotly(
    ggplotly(revenue_plot())
  )

  ## CEO compensation-----------------------------------------------------------
  CEO_plot <- reactive({
    # Testing for the presence of data
    if (sum(!is.na(htmldata()$lederlonn)) == 0 &
      sum(!is.na(htmldata()$leder_annen_godtgjorelse)) == 0) {
      # Printing 'no information' plot if no data
      no_information()
    } else {
      # Creating plot if any data exist
      # Note that the plot will still look empty, but have a legend and scales,
      # if all values are explicit zeroes
      htmldata() |>
        # preparing for pivoting
        select(
          year, lederlonn, leder_annen_godtgjorelse
        ) |>
        # Compensation in ordinary numbers
        mutate(
          lederlonn = lederlonn * 1000,
          leder_annen_godtgjorelse = leder_annen_godtgjorelse * 1000
        ) |>
        # renaming for the purposes of the plot
        rename(
          Year = year,
          Salary = lederlonn,
          `Other compensation` = leder_annen_godtgjorelse
        ) |>
        # pivoting to join the two values to use fill
        pivot_longer(c(Salary, `Other compensation`), names_to = "Compensation") |>
        ggplot(aes(Year, value, fill = Compensation)) +
        geom_col() +
        # using currency scale function to add correct currency code or ?
        currency_scale() +
        # pretty colours
        scale_fill_met_d("Hokusai2", direction = -1) +
        labs(
          x = "", y = "Total compensation", title = "CEO compensation"
        )
    }
  })

  # Creating plotly output
  output$CEO_plotly <- renderPlotly(
    ggplotly(CEO_plot()) |>
      ## ggplot subtitles and captions are lost when using ggplotly(), a
      ## currency warning caption is therefore added in plotly, if needed, using
      ## a separate function
      layout(
        annotations =
          plotly_caption()
      )
  )

  # Reactive functions =========================================================

  ## Currency scale-------------------------------------------------------------
  # Reactive function to check for currency and use currency code or ?
  currency_scale <- reactive({
    scale_y_continuous(
      # Checking for uniform currency code
      labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
        # Using the currency code if uniform
        label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
        # alternative prefix if multiple currencies in the data
        label_dollar(prefix = "? ")
      )
    )
  })

  ## Plotly caption-------------------------------------------------------------
  # Reactive function to create plotly caption (ggplot captions and subtitles
  # are lost in the conversion) (shamelessly stolen from the Internet, with
  # minor modifications).
  # Checks for uniform currency code and prints a warning
  # if there are multiple currency codes.
  plotly_caption <- reactive({
    list(
      x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
        "", "Note: not all values are denominated in the same currency"
      ),
      showarrow = F, xref = "paper", yref = "paper",
      xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
      font = list(size = 11)
    )
  })

  ## scale maximum--------------------------------------------------------------
  # Reactive function to create the maximum value for revenue and expenses plots
  maximum <- reactive({
    # reading the Excel sheet anew
    read_excel(input$excel$datapath) |>
      # removing the firm and sub-category columns
      select(-Firm, -`Sub-category`) |>
      # filtering to retain only revenue and expenses entries
      filter(Category == "Revenue" | Category == "Expenses") |>
      # grouping and summarising to calculate maximum value
      group_by(Year, Category) |>
      summarise(
        sum = sum(Value, na.rm = TRUE)
      ) |>
      # Creating list of sums
      pull() |>
      # selecting the highest value
      max()
  })

  ## Excel data test------------------------------------------------------------
  # Reactive function to test for Excel data quality (basic)
  datatest <- reactive({
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata(), is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }
  })
  # Let the magic/disaster happen (i.e. execute when the user clicks the button)
}) |> bindEvent(input$click)

```{r non-reactive functions, include=FALSE}
# Data reading functions =======================================================

@@

### Test version of function that creates multiple data problems----------------
# read_proff <- function(file, start, end) {
#   read_html(file, encoding = "UTF-8") |>
#     html_node("#inner-frame") |>
#     html_table() |>
#     select( # removing empty column
#       REGNSKAPSPERIODE:all_of(start)
#     ) |> # using all_of()
#     mutate( # removing whitespace
#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
#     ) |>
#     filter( # removing duplicate table - no idea why it's necessary
#       !row_number() > 176
#     ) |>
#     filter( # removing duplicate rows
#       !grepl("Lukk", REGNSKAPSPERIODE)
#     ) |>
#     pivot_longer( # tidying data
#       all_of(end):all_of(start),
#       names_to = "year"
#     ) |> # using all_of()
#     mutate( # changing to real NAs and turning years into numbers
#       value = na_if(value, "-"),
#       year = as.integer(year),
#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
#     ) |>
#     fill( # filling the currency column
#       valutakode,
#       .direction = "updown"
#     ) |>
#     filter( # removing dates and redundant currency
#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
#         REGNSKAPSPERIODE != "Valutakode"
#     ) |>
#     mutate( # removing whitespace in numbers
#       value = str_replace_all(value, "\\s", ""),
#       value = as.numeric(value),
#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
#     ) |>
#     distinct() |> # removing any remaining duplicate rows
#     pivot_wider( # pivoting
#       names_from = REGNSKAPSPERIODE, values_from = value
#     ) |>
#     clean_names() |> # tidying names
#     select(
#       -lederlonn_i_hele_1000,
#       -resultatregnskap_i_hele_1000,
#       -balanseregnskap_i_hele_1000
#     ) |> # removing headings
#     rename( # making it clear what lonn refers to
#       lederlonn = lonn
#     ) |>
#     # Creating new variables for plots
#     mutate(
#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE),
#     ) |>
#     # bad test - REMOVE
#     mutate(
#       valutakode = ifelse(year == 2019, "USD", valutakode),
#       leder_annen_godtgjorelse = NA,
#       lederlonn = NA,
#       arsresultat = NA,
#       sum_egenkapital_og_gjeld = NA
#     ) |>
#     arrange(desc(year)) # arranging by year
# }

@@

@@

# Testing arguments ============================================================
# One or more of the elements of this mutate() operation may be inserted at the
# end of the read_proff() function to create various data problems for some
# quick and dirty testing

# mutate(
#       valutakode = ifelse(year == 2019, "USD", valutakode),
#       leder_annen_godtgjorelse = NA,
#       lederlonn = NA,
#       arsresultat = NA,
#       sum_egenkapital_og_gjeld = NA
#     ) |>
```

>>>>

---
title: "Report"
author: "Torbjørn Skinnemoen Ottersen"
date: '2022-07-24'
output: html_document
params:
  htmlfile: NA
  excelfile: NA
  start: NA
  end: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load(
  shiny, tidyverse, scales, rvest, janitor, plotly, readxl,
  flexdashboard, MetBrewer, fontawesome, htmltools
)
# Using the fontawesome package for more up-to-date (and currently supported)
# Fontawesome icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

# Functions
```{r}
# Data reading functions =======================================================

@@

@@

htmldata <- read_proff(
      params$htmlfile,
      as.character(params$start), as.character(params$end)
    )

exceldata <- widen_excel(params$excelfile)

  ## Excel data test------------------------------------------------------------
  # Reactive function to test for Excel data quality (basic)
datatest <- 
    # let me know if anyone has a better idea for testing
    if (mean(unlist(lapply(exceldata, is.character))) != 0) {
      TRUE
    } else {
      FALSE
    }

@@

## Fundraising ROI------------------------------------------------------------
  # Creating Fundraising ROI value
    # testing for data quality, returning NA if bad
fundraising <- if (datatest == TRUE) {
      NA
    } else {
      exceldata |>
        # most recent year
        filter(year == max(year)) |>
        transmute(
          # Calculating the value as a percentage; removing all other
          KPI_fundraising_num =
            (sum(revenue_private_contributions,
              revenue_membership_fees,
              na.rm = TRUE
            ) - expenses_fundraising) /
              expenses_fundraising * 100
        ) |>
        # just want the single number
        pull() |>
        first()
    } |> KPI_to_string()

```{r data, include=FALSE}

# Key Performance Indicators

### Fundraising ROI: `r fundraising`
